---
title: "Halloween Candy Analysis"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    number_sections: false
---
# Information
**Author: Ross Morton**

> The following notebook is to analyse the halloween_candy_clean.rds file.
The raw data set is a csv for each year (2015-2017) and is stored within the 
/raw_data sub folder.

[Link to raw data](raw_data/)

[Link to cleaning script](cleaning_script/halloween_candy_clean_script.R)

[Link to cleaned data](clean_data/halloween_candy_clean.rds)

> Purpose of this file is to analyse some particular points of the data set.
Primarilay this is:

>1. What is the total number of candy ratings given across the three years. 
(Number of candy ratings, not the number of raters. Don’t count missing values)
2. What was the average age of people who are going out trick or treating?
3. What was the average age of people who are not going trick or treating?
4. For each of joy, despair and meh, which candy bar received the most of these 
ratings?
5. How many people rated Starburst as despair?
+ For the next three questions, count despair as -1, joy as +1, and meh as 0.
6. What was the most popular candy bar by this rating system for each gender in 
the dataset ?
7. What was the most popular candy bar in each year?
8. What was the most popular candy bar by this rating for people in US, Canada, 
UK, and all other countries?


> Installations of the following packages need to be made prior to running the 
script:

+ {tidyverse}
+ {assertr}


# Read in Libraries
```{r}
library(tidyverse)
library(assertr)
```

# Load in data
> Note that the data has been loaded in so that all candy rating columns are
in stored in the one column using pivot longer. Therefore there will be multiple
rows from one individual (this will mess up the age, trick_or_treating and 
country results if not filtered correctly)

```{r}
# Load in the cleaned candy data
candy <- read_rds(file = here::here("clean_data/halloween_candy_clean.rds"))

# View of column names and types
glimpse(candy)
```

## Load in the addtional functions
```{r}
# Function to check the column validity
source(here::here("functions/column_validity_function.R"))

# Function to add the ratings for questions 6 onwards
source(here::here("functions/add_rating_score.R"))
```



# 1. What is the total number of candy ratings given across the three years. 
# (Number of candy ratings, not the number of raters. Don’t count missing values)

Filter the candy ratings column to remove NA's and then count the number of 
rows.

```{r}
candy %>% 
  filter(!is.na(thoughts)) %>% 
  summarise(total_candy_ratings = n())
```


# 2. What was the average age of people who are going out trick or treating?

Select the id, age and trick_or_treat columns- use distinct to remove duplicates
Filter the trick_or_treat for those going out and calculate the average.

```{r}
candy %>% 
  select(id_num, age, trick_or_treat) %>% 
  filter(trick_or_treat == "Yes") %>% 
  distinct() %>% 
  summarise(average_age_guising = mean(age, na.rm = TRUE))
```

# 3. What was the average age of people who are not going trick or treating?

As above but set trick_or_treat to "No".

```{r}
candy %>% 
  select(id_num, age, trick_or_treat) %>% 
  filter(trick_or_treat == "No") %>% 
  distinct() %>% 
  summarise(average_age_not_guising = mean(age, na.rm = TRUE))
```

# 4. For each of joy, despair and meh, which candy bar received the most of these ratings?

Group the candy bars and rating and then count. NA's will show which question
was not answered.

```{r}
candy %>% 
  group_by(thoughts, sweet_name) %>% 
  summarise(rating_num = n()) %>% 
  ungroup(sweet_name) %>% 
  slice_max(rating_num, n=2)
```


# 5. How many people rated Starburst as despair?

Filter for Starburst and the rating as "DESPAIR". Then summarise.

```{r}
candy %>% 
  filter(sweet_name == "starburst"  & thoughts == "DESPAIR") %>% 
  summarise(number_who_hate_starburst = n())
```


>For the next three questions, count despair as -1, joy as +1, and meh as 0.

Create a Function to add the rating system

```{r}
add_rating_score
```
This function also uses the column_validity for checking that the values in the
column are as expected.

```{r}
column_validity
```
Example:

```{r}
candy %>% 
  add_rating_score() %>% 
  select(thoughts, rating) %>% 
  filter(!is.na(thoughts))
```


# 6. What was the most popular candy bar by this rating system for each gender in the dataset ?

Add a column to change the "thoughts" to the rating defined above. Then group
the sweets and complete a sum before slicing for the most popular.

```{r}
candy %>% 
  add_rating_score() %>% 
  group_by(sweet_name) %>% 
  summarise(rating_score = sum(rating, na.rm = TRUE)) %>% 
  slice_max(rating_score, n = 1)
```


# 7. What was the most popular candy bar in each year?

As above initially to change the rating system. Then group by year and sweet 
before slicing.


```{r}
candy %>% 
  add_rating_score() %>%  
  group_by(data_year, sweet_name) %>% 
  summarise(rating_score = sum(rating, na.rm = TRUE)) %>% 
  slice_max(rating_score, n = 1)
```


# 8. What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?


```{r}
candy %>% 
  add_rating_score() %>%  
  group_by(country, sweet_name) %>% 
  summarise(rating_score = sum(rating, na.rm = TRUE)) %>% 
  slice_max(rating_score, n = 1, with_ties = FALSE) %>%
  arrange(match(country,c("United States", "United Kingdom", "Canada")), 
          desc(rating_score))
```

Alternative where grouping the rest of the world together

```{r}
candy %>% 
  add_rating_score() %>%
  mutate(country = if_else(
    (country %in% c("United States", "United Kingdom", "Canada")), country, 
    "Rest of World")) %>% 
  group_by(country, sweet_name) %>% 
  summarise(rating_score = sum(rating, na.rm = TRUE)) %>% 
  slice_max(rating_score, n = 1)
```



# Number who voted for each gender

```{r}
candy %>% 
  group_by(gender) %>% 
  summarise(number_voted = n())
  
```

What gender voted most/least favorably or didn't vote?

```{r}
candy %>% 
  group_by(thoughts, gender) %>% 
  summarise(candy_ratings = n()) %>% 
  arrange(desc(candy_ratings))
```

