---
title: "Seabird Analysis"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    number_sections: false
---
# Information
*Author: Ross Morton*

> The following notebook is to analyse the birds_clean.csv file.
This raw data set is seabirds_raw.xls and is stored within the raw_data sub
folder.

[Link to cleaned data](clean_data/birds_cleaned.csv)

[Link to cleaning script](data_cleaning_scripts/seabirds_clean_script.R)

[Link to raw data](raw_data/seabirds_raw.xls)

> Questions whicch are being analysed
1. Which bird had the most individual sightings?
2. Which bird had the highest total count?
3. Which bird had the highest total count above a latitude of -30?
4. How many different types of birds were only ever seen in groups of 1?
5. How many penguins were seen? (Hint: there are many types of penguin)

> Installations of the following packages need to be made prior to running the 
script:

+ {tidyverse}


# Read in Libraries
```{r}
library(tidyverse)
```


# Read in the file for analysis 
```{r}
birds <- read_csv(here::here("clean_data/birds_cleaned.csv"))
```

# 1. Which bird has the most individual sightings
Most individual sightings- count of _recordings_ for each bird
Filter to obtain the largest number with slice_max
`mutate()` has been used instead of summary to keep the columns. 

>Note that the same birds of different age/sex/plumage have been joined together
in the cleaning process.

```{r}
birds %>% 
  select(contains("species_")) %>% 
  group_by(species_common_name) %>% 
  # Only used mutate to keep the names
  mutate(number_individual_sightings = n()) %>% 
  ungroup() %>% 
  distinct() %>% 
  slice_max(number_individual_sightings, n = 1)
```

# 2. Which bird had the highest total count?

Group the species common name, then sum the `number_birds`

```{r}
birds %>% 
  group_by(species_common_name) %>% 
  summarise(species_scientific_name,
            species_abbreviation,
            total_bird_count = sum(number_birds)) %>% 
  ungroup() %>% 
  distinct() %>% 
  slice_max(total_bird_count, n = 1)
```


However as the number is capped at 99,999- the total number could be higher for
another bird. This value just designates over 100,000 were seen.
Therefore, whilst unlikely the most could be the following instead:

```{r}
birds %>% 
  filter(number_birds >= 99999 & 
           species_common_name != "Short-tailed shearwater") %>% 
  select(-record_id, -latitude, -longitude)
```


# 3. Which bird had the highest total count above a latitude of -30?

Complete similar to before but filter the results for latitude > -30

```{r}
birds %>% 
  filter(latitude > -30) %>% 
  group_by(species_common_name) %>% 
  summarise(species_scientific_name,
            species_abbreviation,
            total_bird_count = sum(number_birds)) %>% 
  ungroup() %>% 
  distinct() %>% 
  slice_max(total_bird_count, n = 1)
```


# 4. How many different types of birds were only ever seen in groups of 1?

Key word is *only ever seen*
Find if the birds have ever been seen with more that one by grouping and using
max.
Then Filter for those with a max of 1. Determine the distinct/unique names
then summarise with the count of the rows.

```{r}
birds %>% 
  group_by(species_common_name) %>% 
  mutate(only_seen_by_itself = max(number_birds) == 1) %>% 
  ungroup() %>% 
  filter(only_seen_by_itself == TRUE) %>% 
  distinct(species_common_name) %>% 
  summarise(number_bird_types = n())

```


# 5. How many penguins were seen? (Hint: there are many types of penguin)

Filter to penguins using `str_detect()` and `regex`. Then sum the number of
penguins.


```{r}
birds %>% 
  filter(str_detect(species_common_name, "[Pp]enguin")) %>% 
  summarise(number_penguins_seen = sum(number_birds))
```


